---
title: "March Madness"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(readr)
mens <- read.csv("march-madness-analytics-2020/2020DataFiles/2020-Mens-Data/MEvents2015.csv")
#mens <-  read.csv("march-madness-analytics-2020/2020DataFiles/2020DataFiles/2020-Mens-Data/MEvents2015.csv")
#mens1 <- read.csv("march-madness-analytics-2020/2020DataFiles/2020DataFiles/2020-Mens-Data/MEvents2016.csv")
#mens2 <- read.csv("march-madness-analytics-2020/2020DataFiles/2020DataFiles/2020-Mens-Data/MEvents2017.csv")
#mens3 <- read.csv("march-madness-analytics-2020/2020DataFiles/2020DataFiles/2020-Mens-Data/MEvents2018.csv")
#mens4 <- read.csv("march-madness-analytics-2020/2020DataFiles/2020DataFiles/2020-Mens-Data/MEvents2019.csv")
#mens <- MEvents2015[complete.cases(MEvents2015),]
#mens1 <- MEvents2016[complete.cases(MEvents2016),]
#mens2 <- MEvents2017[complete.cases(MEvents2017),]
#mens3 <- MEvents2018[complete.cases(MEvents2018),]
#mens4 <- MEvents2019[complete.cases(MEvents2019),]
#MEvents2015
#mens
#mens1
#mens2
#mens3
#mens4
```


```{r}
#Fauziya
#MEvents2015 <- read.csv("march-madness-analytics-2020/2020DataFiles/2020DataFiles/2020-Mens-Data/MEvents2015.csv")
#mens <- MEvents2015[complete.cases(MEvents2015),]

```

####################################################################################
# Wins Box Plot
Here, we study the distribution of the win scores throughout each game season
####################################################################################

```{r}
mens$DayNum<-as.factor(mens$DayNum)
library(ggplot2)
box_plot <- ggplot(mens,aes(x=DayNum, y=WFinalScore)) + geom_boxplot()
box_plot
#boxplot(WFinalScore~DayNum, data = mens)
```


```{r}
mens1$DayNum<-as.factor(mens1$DayNum)
box_plot1 <- ggplot(mens1,aes(x=DayNum, y=WFinalScore)) + geom_boxplot()
box_plot1
```

```{r}
mens2$DayNum<-as.factor(mens2$DayNum)
box_plot2 <- ggplot(mens2,aes(x=DayNum, y=WFinalScore)) + geom_boxplot()
box_plot2
```
```{r}
mens3$DayNum<-as.factor(mens3$DayNum)
box_plot3 <- ggplot(mens3,aes(x=DayNum, y=WFinalScore)) + geom_boxplot()
box_plot3
```

```{r}
mens4$DayNum<-as.factor(mens4$DayNum)
box_plot4 <- ggplot(mens4,aes(x=DayNum, y=WFinalScore)) + geom_boxplot()
box_plot4
```

########################################################################################
# Loss Box plot
Here, we study the distribution of the loose team scores throughout each game season
########################################################################################

```{r}
mens$DayNum<-as.factor(mens$DayNum)
box_plot <- ggplot(mens,aes(x=DayNum, y=LFinalScore)) + geom_boxplot()
box_plot
```

```{r}
mens1$DayNum<-as.factor(mens1$DayNum)
box_plot1 <- ggplot(mens1,aes(x=DayNum, y=LFinalScore)) + geom_boxplot()
box_plot1
```

```{r}
mens2$DayNum<-as.factor(mens2$DayNum)
box_plot2 <- ggplot(mens2,aes(x=DayNum, y=LFinalScore)) + geom_boxplot()
box_plot2
```
```{r}
mens3$DayNum<-as.factor(mens3$DayNum)
box_plot3 <- ggplot(mens3,aes(x=DayNum, y=LFinalScore)) + geom_boxplot()
box_plot3
```

```{r}
mens4$DayNum<-as.factor(mens4$DayNum)
box_plot4 <- ggplot(mens4,aes(x=DayNum, y=LFinalScore)) + geom_boxplot()
box_plot4
```

##############################################################
#barplot
We try to get an overview of typical game events
##############################################################
```{r}
b<-table(mens$EventType)
barplot(b,main = "Event Type Frequency",xlab = "Event Type",col = "red")
```

######################################################################
# Frequency Disttribution
Here, we determine the total number of games won by each team
######################################################################
```{r}
library(dplyr)
m=mens
m <- m %>% 
      group_by(DayNum) %>% 
      distinct(WTeamID,.keep_all = TRUE)
```

# Frequencies
```{r}
y<-table(m$WTeamID)
names(y)[which(y==max(y))]
n<-table(m$LTeamID)
names(n)[which(y==max(n))]
```

```{r}
barplot(y)
barplot(n)
```


#############################################################################################
# Having seen the distribution of wins and losses, we now look closer at a few "important" teams
#############################################################################################
```{r}
z<-sort(y,decreasing = TRUE)
n<-sort(n,decreasing = TRUE)

(v<-z[1:10])
(v1<-z[177:199])
(v2<-z[331:349])
(vec<-z[c(1:3,177:180,331:333)])

barplot(v,col="blue")
barplot(v1,col="blue")
barplot(v2,col="blue")
barplot(vec,col="blue")


```

```{r}

x<-as.data.frame(vec)
names(x)<-c("EventTeamID","NumWins")
d<-as.vector(x$EventTeamID)
MTeams <- read_csv("march-madness-analytics-2020/2020DataFiles/2020DataFiles/2020-Mens-Data/MDataFiles_Stage1/MTeams.csv")
names(MTeams)[1]<-"EventTeamID"
mt<-MTeams[which(MTeams$EventTeamID %in% d),]
nt<-merge(x,mt,by="EventTeamID")
nt1<-merge(nt,m,by="EventTeamID")
barplot(nt$NumWins~nt$TeamName,col="red",ylab = "2015-Number Of Wins",xlab = "TeamName")
```


################################
#2best 2middle 2bottom Teams
################################

```{r}
w<-m[m$WTeamID == c(1246,1458, 1112, 1276, 1287, 1291, 1307, 1123, 1194, 1286), ]

```


```{r}
w[w$LTeamID !=c(1246,1458, 1112, 1276, 1287, 1291, 1307, 1123, 1194, 1286), ]
```


```{r}
p<-c(1246,1458, 1112, 1276, 1287, 1291, 1307, 1123, 1194, 1286)
MPlayers <- read_csv("march-madness-analytics-2020/2020DataFiles/2020DataFiles/2020-Mens-Data/MPlayers.csv")

dat<-merge(mens,MPlayers,by.x=c("EventPlayerID"),
    by.y=c("PlayerID"))

 dat2<-dat[dat$EventTeamID %in% p,]
 #at2
 
 a1<-dat2 %>% 
      group_by(DayNum) %>% 
      distinct(WTeamID,.keep_all = TRUE) 
 
 a2<-dat2 %>% 
      group_by(DayNum) %>% 
      distinct(WTeamID,.keep_all = TRUE) 
 
 
 
a3<-dat2 %>% 
      group_by(DayNum) %>% 
      distinct(WTeamID,.keep_all = TRUE)


a4<-dat2 %>% 
      group_by(DayNum) %>% 
      distinct(WTeamID,.keep_all = TRUE) 

a5<-dat2 %>% 
      group_by(DayNum) %>% 
      distinct(WTeamID,.keep_all = TRUE) 
a6<-dat2 %>% 
      group_by(DayNum) %>% 
      distinct(WTeamID,.keep_all = TRUE) 





#at3
 
 
dat4 <- a1 %>% 
  mutate(WT = ifelse(WTeamID  ==1246 , 1,0))
dat4 <- dat4 %>% 
  mutate(LT = ifelse(LTeamID ==1246, -1,0)) 

dat5 <- a2 %>% 
  mutate(WT = ifelse(WTeamID  ==1458 , 1,0))
dat5 <- dat5 %>% 
  mutate(LT = ifelse(LTeamID ==1458, -1,0))  


dat6 <- a3 %>% 
  mutate(WT = ifelse(WTeamID  ==1276 , 1,0))
dat6 <- dat6 %>% 
  mutate(LT = ifelse(LTeamID ==1276, -1,0))  

dat7 <- a4 %>% 
  mutate(WT = ifelse(WTeamID  ==1287 , 1,0))
dat7 <- dat7 %>% 
  mutate(LT = ifelse(LTeamID ==1287, -1,0))  

 
dat8 <- a5 %>% 
  mutate(WT = ifelse(WTeamID  ==1123 , 1,0))
dat8 <- dat8 %>% 
  mutate(LT = ifelse(LTeamID ==1123, -1,0))  
 
 
  
dat9 <- a6 %>% 
  mutate(WT = ifelse(WTeamID  ==1194 , 1,0))
dat9 <- dat9 %>% 
  mutate(LT = ifelse(LTeamID ==1194, -1,0))  
 
  
 
 

#dat4 <- dat3 %>% 
 # mutate(WT = ifelse(WTeamID  %in% p, 1,0))
#dat4 <- dat4 %>% 
#  mutate(LT = ifelse(LTeamID  %in% p, -1,0)) 
attach(dat4)
dat4<-dat4[order(DayNum),]
dat4$sum <- dat4$WT + dat4$LT
d1<-dat4[which(dat4$WTeamID==1246 | dat4$LTeamID==1246 ),]
 h1<-cumsum(d1$sum)
 d1$h1<-h1
detach(dat4)

attach(dat5)
dat5<-dat5[order(DayNum),]
dat5$sum <- dat5$WT + dat5$LT
 d2<-dat5[which(dat5$WTeamID==1458| dat4$LTeamID==1458),]
 h2<-cumsum(d2$sum)
 d2$h2<-h2
detach(dat5)

attach(dat6)
dat6<-dat6[order(DayNum),]
dat6$sum <- dat6$WT + dat6$LT
d3<-dat6[which(dat6$WTeamID==1276| dat4$LTeamID==1276),]
 h3<-cumsum(d3$sum)
 d3$h3<-h3
detach(dat6)
 

attach(dat7)
dat7<-dat7[order(DayNum),]
dat7$sum <- dat7$WT + dat7$LT
d4<-dat7[which(dat7$WTeamID==1287 | dat4$LTeamID==1287),]
 h4<-cumsum(d4$sum)
 d4$h4<-h4
detach(dat7)

attach(dat8)
dat8<-dat8[order(DayNum),]
dat8$sum <- dat8$WT + dat8$LT
d5<-dat8[which(dat4$WTeamID==1123  | dat4$LTeamID==1123  ),]
 h5<-cumsum(d5$sum)
 d5$h5<-h5
detach(dat8)

attach(dat9)
dat9<-dat9[order(DayNum),]
dat9$sum <- dat9$WT + dat9$LT
 
d6<-dat9[which(dat9$WTeamID==1194)  | dat4$LTeamID==1194 ,]
 h6<-cumsum(d6$sum)
d6$h6<-h6
detach(dat9)


```







```{r}


 
 par(mfrow=c(3,2))
 plot(d1$h1~d1$DayNum,ylab="cumfreq",xlab="DayNum")
 plot(d2$h2~d2$DayNum,ylab="cumfreq",xlab="DayNum")
 plot(d3$h3~d3$DayNum,ylab="cumfreq",xlab="DayNum")
 plot(d4$h4~d4$DayNum,ylab="cumfreq",xlab="DayNum")
 plot(d5$h5~d5$DayNum,ylab="cumfreq",xlab="DayNum")
 plot(d6$h6~d6$DayNum,ylab="cumfreq",xlab="DayNum")

 
 
```












